Closed-loop Neuromorphic Benchmarks

### 2015-10-01 paper/paper.tex
Edits to address reviewer 1

However, \emph{not} using a real physical embodiment for testing is also
problematic.  First and foremost, without an actual real-world task, why
should we have any confidence that the performance on the benchmark is
reflective of the actual usefulness of the neuromorphic hardware?  It is
widely known that simulations of robots (or other physical systems) are
often \emph{much} easier to control and better-behaved than the real thing 
(see \citealt{Jakobi95}; \citealt{Koos2013}).  The
field of robotics is filled with algorithms that work well ``in theory,'' but
fail when run on actual hardware.  We do not want a benchmark that falls into
this trap, giving high scores to hardware that does not turn out to
functional well when deployed in real situations.

Furthermore, neuromorphic hardware has another constraint that severely limits
the utility of simulation.  Typically, when a simulation is
too simple to reflect reality, we add details to the simulation itself.
Incredibly finely detailed simulations can be created, filling in all of the
details needed.  However, accurate modelling of physical systems can very quickly
become \emph{impractical to run in real time}.
This is a fundamental problem, in that neuromorphic
hardware is often tied to real-time interactions, and there is no way to slow down
the hardware to match the simulated environment.  This means that even if
we spent the considerable amount of research effort needed to define a
simulated environment for a closed-loop benchmark, running that simulation 
fast enough to interact with the desired hardware would require more
resources than are available.

We are thus left with a situation where any benchmark we might define for
a closed-loop task will be either impractical for different researchers to
run (if it is physically embodied), or inapplicable to real-world situations
(if it is performed in simulation).  Neither option is particularly appealing.

@@

It is worth noting that a minimal simulation only has to be a
good simulation \emph{for successful behaviour}.  That is, ``if we are evolving 
corridor following behaviour, the dynamics of the simulation might
differ wildly from those of reality if the controller hits a wall or
goes round in circles, but this does
not matter, since the controllers we are interested in transferring
across the reality gap will neither
hit walls nor go round in circles.''  \citep[p.~41]{jakobiThesis}  If the controller is poor, we do not 
need the simulation to be at all accurate in exactly \emph{how} that poor
behaviour is manifest.  We do not need an exact detailed physics model of
the collision between a robot and a wall, or a detailed model of what happens
to a robot arm when it starts oscillating wildly due to a poor control signal.
All we need is for the simulation to be just good enough to indicate that
things have gone wrong, and thus give a low score to that controller.  This
means that, for example, in a minimal simulation of an eight-legged walking robot,
it is not necessary to have a physics simulation that correctly models what
happens when two legs collide with each other.  Rather, if legs collide
with each other, that is an indication that the walking behaviour is very poor.
As long as that result is indicated we can greatly simplify the simulation
by not including all the details necessary to model these complex physical
interactions.

@@

Given the success of this approach for evolutionary robotics, we propose
using minimal simulation for neuromorphic benchmarks.  We would argue that
one important use of a benchmark is \emph{generalization}.  That is, by knowing how well particular
hardware performs on a benchmark, you can reasonably infer how well
that hardware will perform in other situations.  For example, if an image recognition
algorithm performs well on the MNIST hand-written digit recognition
benchmark, this suggests that it may also perform well on
a different recognition task.  Of course, this inference will fail if that
algorithm has been specifically over-fit to that situation.  For
that reason, it is useful to have benchmarks that cover a wide range
of variations on the task.  If the hardware performs well across that
variability, then it is more likely to also work in sufficiently similar new situations.

@@

\subsection{Cheap Robotics}

@@

To keep the physical aspect simple, we recommend cheap, widely-available
components.  This allows a greater chance for other researchers to have
access to the same (or similar) hardware.  For the particular example 
benchmark described in the next section, we use the Lego Mindstorms EV3 kit,
a simple robotics platform available at most toy stores.

It is important to note that there is a theoretical advantage to
using cheap robotics hardware for benchmarking, in addition to the practical
advantages.  In particular, we \emph{do not want benchmarks that rely on
high-speed, high-accuracy devices}.  The purpose of benchmarks
is not to indicate how well this neuromorphic hardware works to control this
particular robot in this task.  Rather, the purpose of a benchmark is to
characterise how well some specific neuromorphic hardware works on a task 
\emph{in general}.  The variability introduced in the minimal simulation means that the hardware
should be able to function across a wide variety of physical embodiments, and
so if we are to choose one particular physical embodiment to test in the
real world, then we should choose one that is not high-precision.
For this reason, we believe using cheap Lego robots is actually more informative
for benchmarking than expensive high-precision robots.\footnote{Of course,
for more complex benchmark tasks we may need sensory and motor capabilities
that are beyond that of a simple Lego robot.}

@@

@@

\subsection{Calibrating the Minimal Simulation via Cheap Robotics}

@@

To run a benchmark using the proposed
minimal simulation approach, there are four main steps: 1. identify
the neuromorphic hardware to be tested; 2. construct the minimal
simulation; 3. determine a metric (e.g. root-mean-squared error; rmse)
to record; 4. specify distributions for any parameters in the
simulation.  We then perform multiple runs of the simulation, each
time choosing different values from the parameter distributions.  We
can then plot how the metric varies as a function of a particular parameter
of interest, or how it compares across different hardware for a
given set of parameter distributions.

@@

All three hardware systems drastically improve performance on this task, as
compared to the non-adaptive controller.

@@

In Figure \ref{fig:analysis_basic}, all three systems perform equally well.
However, on that benchmark all three systems are implementing exactly 500
neurons.  This demonstrates that the
differences in neuron model across that hardware does not significantly
impact performance.  Given that closed-loop models rely on real-time simulation,
it is also important to determine how many neurons each piece of hardware is capable of running in real time.
This is shown in Figure \ref{fig:analysis_compute}.  With the current implementation, the CPU can
run 5200 neurons, the GPU 1500 neurons, and a single SpiNNaker core (of which
there are 18 on one chip) can run 500
neurons.\footnote{The full chip can run much more than 500 neurons; here we use just a single one of the cores on the SpiNNaker chip.}  These values were measured empirically with the current versions
of the reference Nengo implementation, the Nengo OpenCL implementation,
and the Nengo SpiNNaker implementation (as of August 10, 2015).  All other
parameters are as before.

@@

While the primary purpose of this paper is in decribing the benchmarking
methodology, it is also worth noting that these benchmarks indicate
that the neuromorphic learning rule under investigation here is extremely robost.  As
shown in Figure \ref{fig:analysis_neurons_cpu}, even just 500 neurons can consistently
adapt to control a \emph{randomly generated} 15-joint body simulation, and
deal with larger delays and noise than were seen in the example 1-joint
physical embodiment.  Since this learning system is robust across such a
wide range of conditions, and since it is efficiently implementable in a
wide variety of neuromorohic hardware, we feel it is worth further study.  This must include 
both a wider variety of minimal simulation benchmarks and also a few more 
traditional benchmarks.  These traditional
benchmarks would be particular real physical systems (specific robot arms,
for example), but testing on those would only reveal performance on those
particular arms.  As we argued here, benchmarking against a wide variety of 
randomly generated minimal simulation systems is needed to demonstrate the
space of potential situations in which neuromorphic adaptive control performs well.

@@

In addition, other classes of benchmarks could rely on expanded or completely
different minimal simulations.  For example, other physical systems could
be used to calibrate the minimal simulation.  This would lead to other classes
of randomly generated external forces that may be more (or less) difficult
for the neuromorphic system to learn.  If we
identify classes of tasks that we are likely to want to control, we can
create modify those randomly generated forces to ones that are more appropriate
for different tasks.  For example, it
may be of interest to randomly generate $N$-joint arms with random arm
lengths and random masses, and derive (an approximation of) the actual forces 
that would be seen in those situations.

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=18cm]{figures/plot_basic}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:analysis_basic} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems.
     Each system is running 500 neurons.  The hardware do not statistically
     significantly differ, but are all statistically significant improvements 
     over no adaptation (p<0.05).
     Scatterplots show individual runs (with random jitter on the x-axis
     to avoid overlap), the 
     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=18cm]{figures/plot_compute}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:analysis_compute} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems
     in terms of their performance when running as many neurons as they are capable of in real time.  Scatterplots show individual runs (with random
     jitter on the x-axis to avoid overlap), the 
     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  In this case,
 there is no statistical difference between the three hardware systems (p>0.05).}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=18cm]{figures/plot_power}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:analysis_power} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems
     in terms of their performance when running as many neurons as they are capable of per 0.1 Watt of power consumption.  Scatterplots show individual runs (with random jitter on the x-axis to avoid overlap), the 
     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  The SpiNNaker system is a statistically
 significant improvement over the other two (p<0.05).}
\end{figure}

>>>>

However, \emph{not} using a real physical embodiment for testing is also
problematic.  First and foremost, without an actual real-world task, why
should we have any confidence that the performance on the benchmark is
reflective of the actual usefulness of the neuromorphic hardware?  It is
widely known that simulations of robots (or other physical systems) are
often \emph{much} easier to control and better-behaved than the real thing 
(see \citealt{Jakobi95}; \citealt{Koos2013}).  The
field of robotics is filled with algorithms that work well ``in theory,'' but
fail when run on actual hardware.  We do not want a benchmark that falls into
this trap, giving high scores to hardware that does not turn out to
function well when deployed in real situations.

A variety of robotics simulators, such as Webots and Gazebo, already exist
and are intended for evaluating robotics performance.  These are extremely
useful, but have two important limitations.  First, they are generally meant
to evaluate \emph{one particular} robot body, and it is difficult to, for
example, automatically generate a large number of different physical bodies 
to evaluate over.  This means that such a system is good for evaluating a
particular control system for a particular robot body, but is not suitable
for the more general question of how well the control system will work over
a large space of different robot bodies.  Second, these simulators tend
to run slower than real-time.  Typically, when a simulation is
too simple to reflect reality, more details are added to the simulation itself.
Incredibly finely detailed simulations can be created, filling in all of the
details needed.  However, accurate modelling of physical systems can very quickly
become \emph{impractical to run in real time}.
This is a fundamental problem, in that neuromorphic
hardware is often tied to real-time interactions, and there can be 
no way to slow down
the hardware to match the simulated environment.  This means that even if
we spent the considerable amount of research effort needed to define a
simulated environment for a closed-loop benchmark, running that simulation 
fast enough to interact with the desired hardware would require significant
computing resources.  Indeed, one of the major efforts in the Neurorobotics
section of the Human Brain Project is to develop exactly this sort of 
computing infrastructure (\citealt{HBProbot}), with a dedicated supercomputer to run the
physics simulations.  Until this hardware is publically available 
(and until software
is available to create a variety of physical robot models), this approach
is problematic for other researchers.

We are thus left with a situation where any benchmark we might define for
a closed-loop task will be impractical for different researchers to
run (if it is physically embodied), inapplicable to real-world situations
(if it is a simulation that is simple enough to run in real-time), or impossible to connect to
some neuromorphic hardware (if it is a simulation that runs slower than 
real-time).  We thus need a new approach to provide a sharable real-time
simulation that is robust enough that neuromorphic hardware that learns to
deal with the simulation might also be able to deal with reality.

@@

It is worth noting that a minimal simulation only has to be a
good simulation \emph{for successful behaviour}.  That is, ``if we are evolving 
corridor following behaviour, the dynamics of the simulation might
differ wildly from those of reality if the controller hits a wall or
goes round in circles, but this does
not matter, since the controllers we are interested in transferring
across the reality gap will neither
hit walls nor go round in circles.''  \citep[p.~41]{jakobiThesis}  If the controller is poor, we do not 
need the simulation to be at all accurate in exactly \emph{how} that poor
behaviour is manifest.  We do not need an exact detailed physics model of
the collision between a robot and a wall, or a detailed model of what happens
to a robot arm when it starts oscillating wildly due to a poor control signal.
All we need is for the simulation to be just good enough to indicate that
things have gone wrong, and thus give a low score to that controller.  This
means that, for example, in a minimal simulation of an eight-legged walking robot,
it is not necessary to have a physics simulation that correctly models what
happens when two legs collide with each other.  Rather, if legs collide
with each other, that is an indication that the walking behaviour is very poor.
As long as that result is indicated we can greatly simplify the simulation
by not including all the details necessary to model these complex physical
interactions.  This approach was successfully used to develop models of
multi-legged walking robots (\citealt{jakobiThesis}; \citealt{Meyer2003})
and vision-based tracking of a moving object (\citealt{Nolfi2000}).

@@

Although minimal simulation has not previously been used outside the domain of evolutionary robotics, we propose
using minimal simulation for neuromorphic benchmarks.  We would argue that
one important use of a benchmark is \emph{generalization}.  That is, by knowing how well particular
hardware performs on a benchmark, you can reasonably infer how well
that hardware will perform in other situations.  For example, if an image recognition
algorithm performs well on the MNIST hand-written digit recognition
benchmark, this suggests that it may also perform well on
a different recognition task.  Of course, this inference will fail if that
algorithm has been specifically over-fit to that situation.  For
that reason, it is useful to have benchmarks that cover a wide range
of variations on the task.  If the hardware performs well across that
variability, then it is more likely to also work in sufficiently similar new situations.

@@

\subsection{Cost-Effective Robotics}

@@

To keep the physical aspect simple, we recommend cheap, cost-effective, 
widely-available
components.  This allows a greater chance for other researchers to have
access to the same (or similar) hardware.  For the particular example 
benchmark described in the next section, we use the Lego Mindstorms EV3 kit,
a simple robotics platform available at most toy stores.

It is important to note that there is a theoretical advantage to
using simple robotics hardware for benchmarking, in addition to the practical
advantages.  In particular, we \emph{do not want benchmarks that rely on
high-speed, high-accuracy devices}.  The purpose of benchmarks
is not to indicate how well this neuromorphic hardware works to control this
particular robot in this task.  Rather, the purpose of a benchmark is to
characterise how well some specific neuromorphic hardware works on a task 
\emph{in general}.  The variability introduced in the minimal simulation means that the hardware
should be able to function across a wide variety of physical embodiments, and
so if we are to choose one particular physical embodiment to test in the
real world, then we should choose one that is not high-precision.
For this reason, we believe using Lego robots is actually more informative
for benchmarking than expensive high-precision robots.\footnote{Of course,
for more complex benchmark tasks we may need sensory and motor capabilities
that are beyond that of a simple Lego robot.}

@@

It should be noted that, while this algorithm fits well into neuromorphic
hardware, other hardware might be better (in terms of 
accuracy, energy efficiency, cost, or even development time).  Answering
this sort of question is exactly why we need to use a benchmark that can
compare multiple different hardware implementations of this algorithm.  
Furthermore, since some hardware may be better in different situations,
we need a benchmark that has flexible parameters, rather than one that is
based on a single particular physical system.

@@

\subsection{Calibrating the Minimal Simulation via Cost-Effective Robotics}

@@

To run a benchmark using the proposed
minimal simulation approach, there are four main steps: 1. identify
the neuromorphic hardware to be tested; 2. construct the minimal
simulation; 3. determine a metric (e.g. root-mean-squared error; rmse)
to record; 4. specify distributions for any parameters in the
simulation.  We then perform multiple runs of the simulation, each
time choosing different values from the parameter distributions.  For
each run, we reset the hardware to its initial state, so there is no
learning from one run to the next.  This means our metric indicates
how well the system will perform on a single environment, rather than attempting
to use the same learned parameters across different environments.  We
can then plot how the metric varies as a function of a particular parameter
of interest, or how it compares across different hardware for a
given set of parameter distributions.

@@

All three hardware systems drastically improve performance on this task, as
compared to the non-adaptive controller.  Statistical significance was
measured with a two-tailed t-test.

@@

In Figure \ref{fig:analysis_basic}, all three systems perform equally well.
This means that the timing and accuracy differences between the fixed-point
asynchronous SpiNNaker implementation and the floating-point synchronous
CPU/GPU impementations do not affect performance on this task.
However, on that benchmark all three systems are implementing exactly 500
neurons.  This demonstrates that the
differences in neuron model across that hardware does not significantly
impact performance.  Given that closed-loop models rely on real-time simulation,
it is also important to determine how many neurons each piece of hardware is capable of running in real time.
This is shown in Figure \ref{fig:analysis_compute}.  With the current
implementation, including both the Leaky Integrate-and-Fire neuron model
and the learning rule, the CPU can
run 5200 neurons, the GPU 1500 neurons, and a single SpiNNaker core can
run 500 neurons (or 500 $\times$ 16 = 8000 neurons for the whole chip).
These values were measured empirically with the current versions
of the reference Nengo implementation, the Nengo OpenCL implementation,
and the Nengo SpiNNaker implementation (as of August 10, 2015).  All other
parameters are as before.

@@

While the primary purpose of this paper is in describing the benchmarking
methodology, it is also worth noting that these benchmarks indicate
that the neuromorphic learning rule under investigation here is quite robost.  As
shown in Figure \ref{fig:analysis_neurons_cpu}, even just 500 neurons can consistently
adapt to control a \emph{randomly generated} 15-joint body simulation, and
deal with larger delays and noise than were seen in the example 1-joint
physical embodiment.  Since this learning system is robust across such a
wide range of conditions, and since it is efficiently implementable in a
wide variety of neuromorphic hardware, we feel it is worth further study.  This must include 
both a wider variety of minimal simulation benchmarks and also a few more 
traditional benchmarks.  These traditional
benchmarks would be particular real physical systems (specific robot arms,
for example), but testing on those would only reveal performance on those
particular arms.  As we argued here, benchmarking against a wide variety of 
randomly generated minimal simulation systems is needed to demonstrate the
space of potential situations in which neuromorphic adaptive control performs well.

@@

In addition, other classes of benchmarks could rely on expanded or completely
different minimal simulations.  For example, other physical systems could
be used to calibrate the minimal simulation.  This would lead to other classes
of randomly generated external forces that may be more (or less) difficult
for the neuromorphic system to learn.  If we
identify classes of tasks that we are likely to want to control, we can
create modify those randomly generated forces to ones that are more appropriate
for different tasks.  For example, it
may be of interest to randomly generate $N$-joint arms with random arm
lengths and random masses, and derive (an approximation of) the actual forces 
that would be seen in those situations.  In particular, we feel benchmarks
based on the biologically-inspired ``soft-robotics'' systems (e.g. \citealt{Pfeifer2013})
would be particularly appropriate for neural control, given the complexity
involved in generating traditional controllers for them.

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=18cm]{figures/plot_basic}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:analysis_basic} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems.
     Each system is running 500 neurons.  The hardware do not statistically
     significantly differ, but are all statistically significant improvements 
     over no adaptation ($p<0.05$).
     Scatterplots show individual runs (with random jitter on the x-axis
     to avoid overlap), the 
     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=18cm]{figures/plot_compute}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:analysis_compute} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems
     in terms of their performance when running as many neurons as they are capable of in real time.  Scatterplots show individual runs (with random
     jitter on the x-axis to avoid overlap), the 
     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  In this case,
 there is no statistical difference between the three hardware systems ($p>0.05$).}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=18cm]{figures/plot_power}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:analysis_power} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems
     in terms of their performance when running as many neurons as they are capable of per 0.1 Watt of power consumption.  Scatterplots show individual runs (with random jitter on the x-axis to avoid overlap), the 
     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  The SpiNNaker system is a statistically
 significant improvement over the other two ($p<0.05$).}
\end{figure}
